FeatureHashing
==============

Implement feature hashing with R

## Introduction

[Feature hashing](http://en.wikipedia.org/wiki/Feature_hashing) is a popular trick to vectorize text features. 
The package FeatureHashing implements the hashing trick introduced in 
Weinberger et. al. (2009) 
and provides an API similar to model.matrix and Matrix::sparse.model.matrix 
in R. 
The FeatureHashing also supports the splitting of concatenated data during 
the construction of model matrix to save the memory.

### When do I use Feature Hashing?

- I don't need to know the meaning of the cofficients.

    After feature hashing, it is hard to connect the trained cofficients to the original data.

- I cannot convert all data including training data and testing data to a matrix in one time. 

    To make the matrix consistent, I need to synchronize the mapping between the position and the feature. Feature hashing makes my life easier.

- You get a concatenated data in one column.

    Feature Hashing supports the splitting of concatenated data and the user only need to specify the concatenated data and the delimiter in the formula.

## Demo

### Feature Hashing

```{r}
library(FeatureHashing)
# Construct the model matrix. The transposed matrix is returned by default.
m <- hashed.model.matrix(~ ., CO2, 2^6, keep.hashing_mapping = TRUE)
mapping <- as.list(attr(m, "mapping"))
# Print the matrix via dgCMatrix
as(m, "dgCMatrix")[1:20,1:17]

# Check the result of hashing
mapping <- unlist(as.list(attr(m, "mapping")))
mapping %% 2^6 # the 9-th row is conc
# Check the rate of collision
mean(duplicated(mapping %% 2^6))
# The result is CSCMatrix which supports simple subsetting and matrix-vector
# multiplication
# rnorm(2^6) %*% m

# Detail of the hashing
## The main effect is hashed via `hash_h`
all(hash_h(names(mapping)) %% 2^6 == mapping %% 2^6)
## The sign is corrected by `hash_xi`
hash_xi(names(mapping))
## The interaction term is implemented as follow:
m2 <- hashed.model.matrix(~ .^2, CO2, 2^6, keep.hashing_mapping = TRUE)
mapping2 <- unlist(as.list(attr(m2, "mapping")))
mapping2[2] # PlantQn2:uptake
h1 <- mapping2["PlantQn2"]
h2 <- mapping2["uptake"]
library(pack)
hash_h(rawToChar(c(numToRaw(h1, 4), numToRaw(h2, 4)))) # should be mapping2[2]
```

### Concatenated Data

```{r}
# The tag-like feature
data(test.tag)
df <- data.frame(a = test.tag, b = rnorm(length(test.tag)))
head(df)
m <- hashed.model.matrix(~ tag(a, split = ",", type = "existence"):b, df, 2^6,
 keep.hashing_mapping = TRUE)
# The column `a` is splitted by "," and have an interaction with "b":
mapping <- unlist(as.list(attr(m, "mapping")))
names(mapping)
```

