FeatureHashing
==============

Implement feature hashing with R

## Introduction

[Feature hashing](http://en.wikipedia.org/wiki/Feature_hashing), also called as the hashing trick, is a method to
transform features to vector. Without looking the indices up in an
associative array, it applies a hash function to the features and uses their
hash values as indices directly.
  
The package FeatureHashing implements the method in Weinberger et. al. (2009) to transform
a `data.frame` to sparse matrix. The package provides a formula interface similar to model.matrix 
in R and Matrix::sparse.model.matrix in the package Matrix. Splitting of concatenated data, 
check the help of `test.tag` for explanation of concatenated data, during the construction of the model matrix.

### When will I use Feature Hashing?

- I don't need to know the meaning of the cofficients.

    After feature hashing, it is hard to connect the trained cofficients to the original data.

- I cannot convert all data including training data and testing data to a model matrix at once.

    To make the model matrix consistent, I need to synchronize the indices between different conversions. Feature
    hashing automatically synchronize the indices via the hash function.

- I need to handle concatenated data.

    Converting concatenated data to model matrix requires splitting the feature and building some temporal data with `model.matrix`. Feature Hashing supports the splitting of the concatenated data on the fly with simple formula interface. The user does not need to call `strsplit` and spend a lot of memory for temporal data.

## Demo

### Concatenated Data

```{r}
library(methods)
library(FeatureHashing)

# The tag-like feature
data(test.tag)
df <- data.frame(a = test.tag, b = rnorm(length(test.tag)))
head(df)
m <- hashed.model.matrix(~ tag(a, split = ",", type = "existence"):b, df, 2^6,
 keep.hashing_mapping = TRUE)
# The column `a` is splitted by "," and have an interaction with "b":
mapping <- unlist(as.list(attr(m, "mapping")))
names(mapping)
```

### Feature Hashing

```{r}
# Construct the model matrix. The transposed matrix is returned by default.
m <- hashed.model.matrix(~ ., CO2, 2^6, keep.hashing_mapping = TRUE)
mapping <- as.list(attr(m, "mapping"))
# Print the matrix via dgCMatrix
as(m, "dgCMatrix")[1:20,1:17]

# Check the result of hashing
mapping <- unlist(as.list(attr(m, "mapping")))
mapping %% 2^6 # the 9-th row is conc
# Check the rate of collision
mean(duplicated(mapping %% 2^6))
# The result is CSCMatrix which supports simple subsetting and matrix-vector
# multiplication
# rnorm(2^6) %*% m

# Detail of the hashing
## The main effect is hashed via `hash_h`
all(hash_h(names(mapping)) %% 2^6 == mapping %% 2^6)
## The sign is corrected by `hash_xi`
hash_xi(names(mapping))
## The interaction term is implemented as follow:
m2 <- hashed.model.matrix(~ .^2, CO2, 2^6, keep.hashing_mapping = TRUE)
mapping2 <- unlist(as.list(attr(m2, "mapping")))
mapping2[2] # PlantQn2:uptake
h1 <- mapping2["PlantQn2"]
h2 <- mapping2["uptake"]
library(pack)
hash_h(rawToChar(c(numToRaw(h1, 4), numToRaw(h2, 4)))) # should be mapping2[2]
```


